---
title: "Assessment data - processing"
author: "Molly Wilson"
date: "11/12/2020"
output: html_document
---

```{r, message = F, warning = F, echo = F}
library(startR)
library(tidyverse)
library(here) # for accessing files with directory
library(readxl) # for reading indv. sheets within excel files
library(janitor) # for cleaning names so they have consistent capitalization etc.
library(viridis)
library(ggforce)

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE) # this sets the preferences for each code chunk so that errors/messages/warnings don't get displayed in the knit rmarkdown
```

```{r}
# Import data

fish <- read_excel(here("data", "MRC_fish.xlsx"), sheet = "mrc.fish") %>% clean_names() %>%
  select(-c(date, time, depth_m, surveyor)) %>% # not using these variables currently, so nice to streamline
  filter(!is.na(site)) %>% # remove any incomplete rows at the end of the data
  uncount(number) # expand to replicate rows if multiple fish were recorded

rugosity <- read_excel(here("data", "MRC_data.xlsx"), sheet = "mrc.rugosity") %>% clean_names() %>%
   select(-c(date, time, depth_m, surveyor))

benthic <- read_excel(here("data", "MRC_benthic.xlsx"), sheet = "benthic") %>% clean_names() %>%
  filter(!is.na(site)) %>%
  select(-c(date, surveyor, notes))

benthic2 <- read_excel(here("data", "MRC_benthic_V5.xlsx"), sheet = "benthic") %>% clean_names() %>%
  filter(!is.na(site)) %>%
  select(-c(date, surveyor, notes))

detach(janitor)
```

Target plots:

Fish:
  - Biomass of functional groups per site, maybe total biomass/herbivore biomass () regional average pristine
  - Size of terminal phases compared to regional average 
  
Benthic:
  - broad categories by site (on a map?)
  - CORALS: abundance of different species/genus/shape 
  - stressed vs. unstressed corals by species/genus 
  - across algae - turfs/macroalgae (one with only MA) by type, what is palatable and to what
  - sedimentation - TA by type - by site
  - (average dictyota height by site)
 
Rugosity:
  - by site, compared to regional/pristine average
  
Urchins:
  - density by site compared to historical reference 

## Benthic analyses

### Percent cover of broad categories for overview graph

```{r}
# Calculating percent cover per summary categories

# First, we need to add a few more classification variables for calculating percent cover and for future graphs
benthic <- benthic %>%
  mutate(benthic_class = case_when(category_code %in% c("LC") ~ "Live hard corals",
                                  category_code %in% c("PDC", "LCD", "PLC") ~ "Stressed hard corals",
                                  category_code %in% c("MA", "FMA") ~ "Macroalgae",
                                  category_code %in% c("TA") ~ "Turf algae",
                                  category_code %in% c("CCA", "ND CCA", "CCA - FB", "PEY") ~ "CCA",
                                  category_code %in% c("OCTO", "SPON") ~ "Soft coral and sponges",
                                  category_code %in% c("AINV", "CYAN") ~ "AINV + CYAN",
                                  category_code %in% c("SAND", "HOLE", "SG", "PAVE") ~ "Other substratum"
                                    ), # manually adding summary categories for simplified graphs
         direction = case_when (category_code %in% c("LC", "CCA", "ND CCA", "CCA - FB", "PEY", "PDC", "LCD", "PLC") ~ "Corals & facilitators",
                                category_code %in% c("MA", "FMA", "TA", "OCTO", "SPON", "AINV", "CYAN") ~ "Inhibitors"), # classifying categories as good or bad; codes not included will come up as "NA"
         av_sub_yn = if_else(category_code %in% c("SAND", "HOLE", "SG", "PAVE"), "no", "yes") # if its in this list, considered not available substrate, all other codes are considered "yes"
         )  %>%
  filter(!(site == "Rickett's Bay" & ((transect == 1 & meter == 9) | (transect == 2 & meter == 9)))) # both of these meters were only sand and/or seagrass, and were causing issues later in the analysis

# first we need to calculate percent cover per each meter
benthic.pc.m.raw <- benthic %>%
  filter(av_sub_yn == "yes") %>% # only looking at what is considered available substrate (no sand, seagrass, etc.)
  group_by(site, transect, meter) %>%
  mutate(n_pts = n()) %>% # showing total number of points per meter that are considered available substrate
  ungroup() %>%
  group_by(site, transect, meter, n_pts, benthic_class) %>%
  summarize(pc.m = 100*n()/n_pts) %>% # n() counts the number of entries within a given group (i.e. points per species here)
  ungroup() %>%
  distinct() %>%
  select(-n_pts)

# "expand" a dataframe so that it contains all species categories per site/transect/meter, and then join the full dataset so that any meters where a species was absent it will show up as 0 (as opposed to just not having an entry at all)
benthic.pc.m <- benthic %>%
  expand(nesting(site, transect, meter), benthic_class) %>%
  left_join(benthic.pc.m.raw, by = c("site", "transect", "meter", "benthic_class")) %>%
  mutate(pc.m = if_else(is.na(pc.m), 0, pc.m))

# calculating percent cover over summary benthic classes
benthic.pc.class <- benthic.pc.m %>%
  group_by(site, transect, benthic_class) %>%
  summarize(pc.t = mean(pc.m)) %>%
  ungroup() %>%
  group_by(site, benthic_class) %>%
  summarize(pc = mean(pc.t)
            #se = sd(pc.t)/sqrt(n.tran)
            ) %>%
  left_join(benthic %>%
              select(benthic_class, direction) %>%
              distinct(),
            by = "benthic_class") %>%
  filter(benthic_class != "Other substratum") %>%
  arrange(direction)

# quick check to make sure everything adds up to 100% at the meter and site level
test.m1 <- benthic.pc.m %>%
  group_by(site, transect, meter) %>%
  summarize(total = sum(pc.m))
test.site <- benthic.pc.class %>%
  group_by(site) %>%
  summarize(total = sum(pc))

ggplot(benthic.pc.class, aes(x = benthic_class, y = pc, fill = direction)) +
  geom_col() +
  scale_fill_manual(values = c("aquamarine3", "coral2")) +
  facet_grid(. ~ site) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Benthic class", y = "Percent cover")

ggplot(data = benthic.pc.class %>%
              mutate(pc = if_else(direction == "Inhibitors", -pc, pc)), 
      aes(x = benthic_class, y = pc, fill = direction)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values = c("aquamarine3", "coral2")) +
  facet_grid(site ~ .) +
  labs(x = "Benthic class", y = "Percent cover") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


benthic.pc.class <- benthic.pc.class %>% 
  arrange(desc(benthic_class)) %>%
  mutate(ypos = cumsum(pc)- 0.5*pc )
ggplot(benthic.pc.class, aes(x = "benthic_class", y = pc, fill = benthic_class)) + 
  geom_bar(width = 1, stat = "identity", color = "black") +
  coord_polar("y", start=0) +
  facet_wrap(vars(site), nrow = 2) +
  #geom_text(aes(y = ypos, label = benthic_class), color = "white", size=2) +
  theme_void() +
  theme(legend.position = "bottom")

# benthic.pc.class <- benthic.pc.class %>% 
#   mutate(end = 2 * pi * cumsum(pc)/sum(pc),
#          start = lag(end, default = 0),
#          middle = 0.5 * (start + end),
#          hjust = ifelse(middle > pi, 1, 0),
#          vjust = ifelse(middle < pi/2 | middle > 3 * pi/2, 0, 1))
# ggplot(benthic.pc.class) + 
#   geom_arc_bar(aes(x0 = 0, y0 = 0, r0 = 0, r = 1,
#                    start = start, end = end, fill = benthic_class)) +
#   geom_text(aes(x = 1.05 * sin(middle), y = 1.05 * cos(middle), label = benthic_class,
#                 hjust = hjust, vjust = vjust), size = 3) +
#   coord_fixed() +
#   scale_x_continuous(limits = c(-1.8, 1.8),  # Adjust so labels are not cut off
#                      name = "", breaks = NULL, labels = NULL) +
#   scale_y_continuous(limits = c(-1.8, 1.8),      # Adjust so labels are not cut off
#                      name = "", breaks = NULL, labels = NULL) +
#   facet_wrap(vars(site), nrow = 2) +
#   theme_void() +
#   theme(legend.position = "none")
```

### Relative abundance of coral species - should this be average of percent cover, 

```{r}
benthic.pc.spp.m.raw <- benthic %>%
  filter(av_sub_yn == "yes") %>% # only looking at what is considered available substrate (no sand, seagrass, etc.)
  group_by(site, transect, meter) %>%
  mutate(n_pts = n()) %>% # showing total number of points per meter that are considered available substrate
  ungroup() %>%
  group_by(site, transect, meter, species_code) %>%
  summarize(pc.spp.m = 100*n()/n_pts) %>%
  ungroup() %>%
  distinct()

benthic.pc.spp.m <- benthic %>%
  expand(nesting(site, transect, meter), species_code) %>%
  left_join(benthic.pc.spp.m.raw, by = c("site", "transect", "meter", "species_code")) %>%
  mutate(pc.spp.m = if_else(is.na(pc.spp.m), 0, pc.spp.m)) %>%
  # left_join(benthic %>% 
  #             select(species_code, species, genus = type) %>%
  #             distinct(),
  #           by = "species_code")

temp <-benthic %>% 
              select(species_code, species, genus = type) %>%
              distinct()
test.m1 <- benthic.pc.spp.m %>%
  group_by(site, transect, meter) %>%
  summarize(total = sum(pc.spp.m))

corals.pc.spp <- corals.exp %>%
  group_by(site, transect, genus, species_code, species) %>%
  summarize(pc.t = mean(pc.m)) %>%
  group_by(site, genus, species_code, species) %>%
  summarize(pc = mean(pc.t))
            #se = sd(pc.t)/sqrt(max(transect)) # need to fix SE 

corals.pc.spp.tot <- corals.pc.spp %>%
  group_by(genus, species_code, species) %>%
  summarize(pc.tot = mean(pc)) %>%
  arrange(pc.tot)

ggplot(corals.pc.spp.tot, aes(x = reorder(species, pc.tot), y = pc.tot, fill = genus)) +
  geom_col() +
  coord_flip() +
  labs(x = "", y = "Mean percent cover") +
  scale_color_viridis() +
  theme_bw()
  
```
## Fish

### Biomass by site by species, family, and functional group

```{r}
fish.bm.spp.t.raw <- fish %>%
  filter(biomass > 0) %>% # only looking at what is considered available substrate (no sand, seagrass, etc.)
  group_by(site, transect, species_code) %>%
  summarize(bm.spp.t = sum(biomass)/1.2)

fish.bm.spp.t <- fish %>%
  expand(site, transect, species_code) %>%
  left_join(fish.bm.spp.t.raw, by = c("site", "transect", "species_code")) %>%
  mutate(bm.spp.t = if_else(is.na(bm.spp.t), 0, bm.spp.t)) %>%
  left_join(fish %>%
              select(species_code, species_name, common_name, family, functional_group) %>%
              unique(),
                      by = "species_code")

fish.bm.spp.site <- fish.bm.spp.t %>%
  group_by(site, species_code, species_name, common_name) %>%
  summarize(n.t = n(),
            bm = mean(bm.spp.t),
            se = sd(bm.spp.t)/sqrt(n.t))

fish.bm.fam.site <- fish.bm.spp.t %>%
  group_by(site, family, transect) %>%
  summarize(bm.fam.t = sum(bm.spp.t)) %>%
  ungroup() %>%
  group_by(site, family) %>%
  summarize(n.t = n(),
            bm = mean(bm.fam.t),
            se = sd(bm.fam.t)/sqrt(n.t))

fish.bm.func.site <- fish.bm.spp.t %>%
  group_by(site, functional_group, transect) %>%
  summarize(bm.func.t = sum(bm.spp.t)) %>%
  ungroup() %>%
  group_by(site, functional_group) %>%
  summarize(n.t = n(),
            bm = mean(bm.func.t),
            se = sd(bm.func.t)/sqrt(n.t))

ggplot(fish.bm.func.site, aes(x = functional_group, y = bm, group = functional_group, fill = functional_group)) +
  geom_col() +
  facet_grid(. ~ site) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
  
```


